---
title: "Titanic"
author: "Frederic TWAHIRWA"
date: "2 février 2017"
output: html_document
---

```{r}
# clean up the memory of the current R session
rm(list=ls(all=TRUE))
```

```{r}
# load packages
library(dplyr)
library(tidyr)
```

Loading a csv file in RStudio into a data frame; Here I am using read.csv instead of read.table which expects an element in each column 
All missing (empty) values are replaced  by the string 'NA' (args ) na.strings=c("", "NA") in the read.csv() function
```{r}
# load dat set in tamporaly data set "titanicf"
titanicf<-read.csv('titanic_original.csv', header = T, sep=';', dec=",", na.strings=c("", "NA"))

# replace missing value in embarked column with "s"
titanicf$embarked[which(is.na(titanicf$embarked))]<-"S"
```

```{r}
# the summary() function provides for each variable in your dataset the minimum, mean, maximum and quartiles
summary(titanicf$age)
```

```{r}
# replace the values in the Age column are missing by the mean (29.88)
# this missing value could be repaced by the median value  (28.00)

age_mean<-mean(titanicf$age, na.rm = TRUE)
age_median<-median(titanicf$age, na.rm = TRUE)

titanicf$age[which(is.na(titanicf$age))]<-mean(titanicf$age, na.rm = TRUE)
```
In cabin column, it seems that there is a lot of data not recorded in this field, special in second and third class
may be this column reflects who booked with cabins.

```{r}
#create a new column, in this case I am using "dplyr" with mutate, as.numeric and !is.na allow to coerce the TRUE logic variable in 0 numeric and FALSe logic variable in 1 numeric
titanicf<-titanicf%>% 
        mutate(has_cabin_number=as.numeric(!is.na(titanicf$cabin)))
```

```{r}
# write a csv file on my working directory
write.csv(titanicf, file = "titanic_clean.csv")

```
